Conformal Prediction for Students' Grades in a Course Recommender System

### 2019-05-22 Conference/COPA2019/paper/draft.tex
draft submitted to conference for review

@@

% The following command is just for this sample document:
\newcommand{\cs}[1]{\texttt{\char`\\#1}}

 % Define an unnumbered theorem just for this sample document:
\theorembodyfont{\upshape}
\theoremheaderfont{\scshape}
\theorempostheader{:}
\theoremsep{\newline}
\newtheorem*{note}{Note}

@@

% start article
% \titlebreak
% \footnote{}
% \textsf
% \bfseries

\title[Conformal Prediction for Students' Grades]{Conformal Prediction for Students' Grades in a Course Recommender System\titletag{\thanks{XXX}}}

@@

 % Authors with different addresses:
  \author{\Name{Rapha\"{e}l Morsomme\nametag{\thanks{with a note}}} \Email{raphael.morsomme@maastrichtuniversity.nl}\\
   \addr Zwingelput 4, 6211 KH Maastricht, the Netherlands	
   \AND
   \Name{Evgueni Smirnov} \Email{smirnov@maastrichtuniversity.nl}\\
   \addr DKE's address, 6211 KH Maastricht, the Netherlands}

@@

\begin{abstract}
Course selection can be challenging for student of Liberal Arts programs. In particular, due to the highly personalized curricula of liberal arts students, it is often difficult to assess whether or not a certain course is too advanced given their academic background. To assist students of the liberal arts program of the University College Maastricht, Morsomme and Vazquez (2019) developed a course recommender system that suggests courses whose content matches the student's academic interests, and issues warnings for courses that it deems too advanced. To accomplish the latter, we fit a sparse regression model for grade. In this paper, we use the Inductive Confidence Machine (Papadopoulos, 2002) to obtain predictive regions for future grades. We compare the results obtained with the standard regression nonconformity measure and the normalized nonconformity measure. We find that both nonconformity measures produce valid predictive regions and that the width of the predictive regions is correlated with the accuracy of the regression model.
\end{abstract}

\begin{keywords} Conformal Prediction, Recommender System, Grade Prediction, Lasso, Education \end{keywords}

\todo{Rapha\"{e}l: make all figures black and white}

\todo{question: make flowchart with two pillars (course recommendation and warnings) instead of input/model/intermediate results/output?}

\todo{connect with bibtex bibliography from citavi}

@@

difficult to assess if student ready. 
Morsomme and Vazquez include warnings based on grade prediction to help student identify course too advanced for them. Provide warning if predicted grade is a fail.

In this paper, we present an applictaion of the ICM to provide students with prediction regions instead of warnings based on a point estimate for the grade. The advantage of PR over point estimates is that it improves info position of student, tehreby enabling them to make better-informed course selection. We decide to use conformal prediction prediction to tailor the PR to the students and to use an ICM to alleviate the computational costs of transductive.

Liberal Arts program are often characterized by their open curriculum which allows student to tailor their own study program to their academic objectives. These highly personalized curricula make it difficult for students, academic advisors and course coordinators to assess whether a particular course is too advanced given the student's current academic background or whether she/he has acquired the prerequisite skills via a combination of courses. To alleviate this problem, Morsomme and Vazquez (2019) developed a course recommender system which suggest courses whose content matches the academic interests of the student and issues warnings for courses that it deems too advanced given the student's academic background. To accomplish the latter, they fit a sparse regression model to each course which takes as input a model of the student indicating the skills that she/he has acquired in previous courses and her/his academic performance in different academic disciplines in order to predict grades. Warnings are issued when the system predicts a fail grade.

This paper presents the application of an Inductive Confidence Machine (ICM) (Papadopoulos, 2002) to complement the existing course recommender system of [ref: Morsomme and Vazquez, 2019] with predictive regions for future grades tailored to the student.

To improve student info position, we complement point estimates predictions with region predictions. To tailor the prediction to each student, we use conformal prediction.

This way, the information position of students is improved; instead of providing students with a binary advice, i.e. a warning if the predicted grade is a fail and nothing if it is a passing grade, students are provided with a predictive regions indicating a range of grades they are likely to receive in the courses that they are, for instance, considering for the following term particular course.

We present previous research on grade prediction in \sectionref{sec:related} and the data in \sectionref{sec:data}
We introduce the methodology related to the course recommender system and the ICM in \sectionref{sec:methodology} and present the results in \sectionref{sec:results}

\section{Related Work (Evgueni)}
\label{sec:related}

Conclusion of the section: no previous research has applied the conformal prediction framework to students' grade prediction.

\section{Data}
\label{sec:data}

We employ two sets of data: student data and course data.

The student data consists of anonymized course enrollment information. We use the transcripts of the 2,526 students of the liberal arts program between 2008 and 2019 with a total of 79,245 course enrollments. We exclude enrollments with a missing grade which indicates that the student either dropped the course or fail the attendance requirement. In the latter case, the data set contains an observation corresponding to the resit. \tableref{tab:student-data} presents the student data. Each row contains an anonymized student ID, a course ID, a year and semester, and the obtained grade.

The course data consists of the 2018-2019 course catalogues of five departments of Maastricht University: European Studies, University College Maastricht, University College Venlo, Psychology and Science Program. These catalogues contain a one-page description of 490 courses. \tableref{tab:course-data} presents the textual data in the tidy format with one row per document-term [ref: Wickham 2014]. We process the data following common cleaning procedures [ref: Meyer, 2008]: we tokenize the individual terms, stem them with the Hunspell dictionary and remove common stop words, numbers between 1 and 1,000, and terms occurring less than 3 times in the data set.

\begin{table}[hbtp]
	\floatconts
	{tab:student-data}
	{\caption{Example of Student Data}}	
	{\begin{tabular}{ccccc}
		\toprule
		\bfseries Student ID &\bfseries Course ID &\bfseries Academic Year &\bfseries Period &\bfseries Grade\\
		\midrule
		44940 & CAP3000 & 2009-2010 & 4 & 8.8\\
		37490 & SSC2037 & 2009-2010 & 4 & 8.4\\
		71216 & HUM1003 & 2010-2011 & 4 & 6.8\\
		44212 & SSC2049 & 2010-2011 & 2 & 8.4\\
		85930 & SSC2043 & 2011-2012 & 1 & 4.3\\
		\addlinespace
		14492 & COR1004 & 2012-2013 & 2 & 8.5\\
		34750 & HUM2049 & 2013-2014 & 5 & 6.0\\
		32316 & SSC1001 & 2013-2014 & 1 & 8.5\\
		22092 & SCI1009 & 2014-2015 & 1 & 6.4\\
		19512 & COR1004 & 2016-2017 & 5 & 7.0\\
		\bottomrule
	\end{tabular}}	
\end{table}

\begin{table}[hbtp]
	\floatconts
	{tab:course-data}
	{\caption{Example of Course Data}}	
	{\begin{tabular}{cccc}
			\toprule
			\bfseries Course ID &\bfseries Course Title &\bfseries Department &\bfseries Term\\
			\midrule
			HUM3034 & World History & UCM & understand\\
			HUM3034 & World History & UCM & major\\
			HUM3034 & World History & UCM & issue\\
			HUM3034 & World History & UCM & episode\\
			HUM3034 & World History & UCM & shape\\
			\addlinespace
			HUM3034 & World History & UCM & history\\
			HUM3034 & World History & UCM & mankind\\
			HUM3034 & World History & UCM & focus\\
			HUM3034 & World History & UCM & theme\\
			HUM3034 & World History & UCM & topic\\
			\bottomrule
	\end{tabular}}	
\end{table}

\section{Methodology}
\label{sec:methodology}

\subsection{The Course Recommender System}
\label{sec:crs}

\subsubsection{Overview}
\label{sec:overview}
The course recommender system has two goals: suggesting courses whose content matches the student's academic interests and help the student assess whether she/he has acquired the necessary knowledge to succeed in the course she /he is considering for the following term. Although the former is not of direct interest for this paper, we briefly describe

\begin{figure}[htbp]
	\floatconts
	{fig:flowchart}
	{\caption{Flowchart of the Course Recommender System}}
	{\includegraphics[width=0.5\linewidth]{figures/flowchart}}
\end{figure}

\subsubsection{Topic Model}
\label{sec:tm}
We start by fitting a topic model to the course data using the Latent Dirichlet Allocation generative model and the Gibbs sampling algorithm (ref: Blei, 2003; Phan, 2008). The LDA model views a topic as a mixture of words and a document as a mixture of topics. More precisely, topics are conceptualize as a probability distribution over the vocabulary of the data, and document as a set of words, each drawn from a probability distribution over topics specific to that document. The term \textit{Dirichlet} comes from the fact that the word distribution $\beta_{t}$ for a topic $t$ is generated from a Dirichlet distribution $\beta_{t} \sim Dirichlet(\delta)$ and the topic distribution $\theta_{d}$ for a document $d$ is generated from a Dirichlet distribution $\theta_{d} \sim Dirichlet(\alpha)$ where $\delta$ and $\alpha$ act as hyper-parameters determining how concentrated the distributions of words in topics and the distributions of topics in documents are.

We follow Phan (ref: 2008) who used a Gibbs sampler to learn the distributions $\beta$ and $\theta$. The obtained topic model consists of a term distribution for each topic and a topic distribution for each course description. \figureref{fig:gamma} shows the importance of the main topics of the core course COR1004 Political Philosophy. \figureref{fig:beta} indicates that the two main topics of the course (topics 4 and 19) respectively correspond to the themes of international politics and philosophy.

The number of topics $T$ needs to be specified \textit{a-priori}. We trained 30 models with $K=5,10,\cdots,150$. As suggested by Griffiths and Steyvers (ref: 2008), we set $\alpha=50/K$ and $\delta=0.1$. We run 6,000 iterations of the Gibbs sampler with a burn in of 1,000 iterations and sample every 100 iterations. To prevent being stuck in a local optimum, we use 10 random initialization and keep the best model with respect to the log-likelihood. We choose the number of topics $K$ that yields the model with the maximum likelihood. Figure X shows that the model with 65 topics has the maximum likelihood. To increase the quality of the model, we retrain the model with 16,000 iterations of the the Gibbs sampler, a burn in of 2,000 iterations and 20 random starts and the other parameters at their original values.

\begin{figure}[htbp]
	\floatconts
	{fig:gamma}
	{\caption{Topic Distribution in the course COR1004 Political Philosophy.}}
	{\includegraphics[width=0.5\linewidth]{figures/gamma-COR1004}}
\end{figure}

\begin{figure}[htbp]
	\floatconts
	{fig:beta}
	{\caption{Word Distribution in the main two topics of COR1004 Political Philosophy. Topic 4 corresponds to international politics and Topic 19 to philosophy.}}
	{
		\subfigure[Topic 4]{\label{fig:beta-4}
			\includegraphics[width=0.2\linewidth]{figures/beta-4}}
		\qquad
		\subfigure[Topic 19]{\label{fig:beta-19}%
			\includegraphics[width=0.2\linewidth]{figures/beta-19}}
	}
\end{figure}

\begin{figure}[htbp]
	\floatconts
	{fig:ll}
	{\caption{Maximum Likelihood Model Selection: we choose the model with 65 topics.}}
	{\includegraphics[width=0.5\linewidth]{figures/model-selection}}
\end{figure}

\subsubsection{Student Model}
\label{sec:sm}
The student model consists of two elements: academic performance and level of topic-specific expertise. Academic performance corresponds to the student's general GPA as well as her/his GPA in humanities, natural sciences, social sciences, skills and projects. These are derived from the student's transcript in a straightforward way. Topic-specific expertise corresponds to the knowledge that the student has acquired in each of the topics present in the topic model in previous courses. We posit that when students takes a course they acquire knowledge about its content and that the amount of knowledge they acquire is proportional to their grade. More precisely, we assume that students who obtain 10/10 in a course acquire all the knowledge related to its content and that those who obtain 5/10 only acquire half of it. The content of a course is determined by its topic distribution in the topic model and the grade are retrieved from the student's transcript. Hence if a student has taken $n$ courses and  $g_{i}$ corresponds to her/his grade in  course $i$, for $i = 1, \cdots, n$, then the expertise of the student $e_t$ in topic $t$ simply corresponds to $$e_t = \sum_{i= 1}^{n} g_i \gamma_{i,t}$$ where $\gamma_{i,t}$ corresponds to the importance of topic $t$ in course $i$ in the topic model. \figureref{fig:toy-expertise} presents a toy example of the contribution of individual courses to a student's expertise in 5 topics.

\begin{table}[htbp]
	\floatconts
	{tab:toy}
	{\caption{Toy Example of Topic Expertise}}
	{%
		\subtable[Topic Distribution $\gamma$]{%
			\label{tab:toy-gamma}%
			\begin{tabular}{cccccc}
				\toprule
				\bfseries Course & \bfseries Topic 1 & \bfseries Topic 2 & \bfseries Topic 3 & \bfseries Topic 4 & \bfseries Topic 5\\
				\midrule
				Course 1 & 0.0 & 0.7 & 0.2 & 0.1 & 0.0\\
				Course 2 & 0.2 & 0.2 & 0.2 & 0.2 & 0.2\\
				Course 3 & 0.0 & 0.4 & 0.2 & 0.1 & 0.2\\
				\bottomrule
			\end{tabular}
		}\qquad
		\subtable[Transcript]{%
			\label{tab:toy-grade}%
			\begin{tabular}{cc}
				\toprule
				\bfseries Course & \bfseries Grade\\
				\midrule
				Course 1 & 6/10\\
				Course 2 & 9/10\\
				Course 3 & 2.5/10\\
				\bottomrule
			\end{tabular}
		}\qquad
			\subtable[Course Contribution towards Topic Expertise]{%
			\label{tab:toy-contribution}%
			\begin{tabular}{cccccc}
				\toprule
				\bfseries Course & \bfseries Topic 1 & \bfseries Topic 2 & \bfseries Topic 3 & \bfseries Topic 4 & \bfseries Topic 5\\
				\midrule
				Course 1 & 0.00 & 0.42 & 0.12 & 0.060 & 0.00\\
				Course 2 & 0.18 & 0.18 & 0.18 & 0.180 & 0.18\\
				Course 3 & 0.00 & 0.10 & 0.05 & 0.025 & 0.05\\
				\bottomrule
			\end{tabular}
		}
	}
\end{table}

\begin{figure}[htbp]
	\floatconts
	{fig:toy-expertise}
	{\caption{Contribution of Individual Courses towards a Student's Topic Expertise Profile.}}
	{\includegraphics[width=0.5\linewidth]{figures/toy-expertise}}
\end{figure}

\subsubsection{Underlying Algorithm: Sparse Regression Model}
\label{sec:sm}

We fit a sparse linear regression model for grade prediction to each of the 132 courses currently offered at the college with at least 20 student enrollments since 2008. The input consists of the 71 variables present in the student model at the start of the course: her/his 6 GPA's  (1 general and 5 discipline-specific) and 65 levels of topic-specific expertise. Since the number of predictors is relatively large, we regularize the models with the lasso penalty (ref: Tibshirany, 1998). For each course, we use 10-fold cross-validation to find the value of the lasso tuning parameter $\lambda$ that minimizes the CV mean absolute error (mae), a more robust loss function than the squared absolute error (ref: els, 2008). \figureref{fig:mae-cv} presents the distribution of the CV mae for the 132 prediction models. The model for the course \textit{PRO2004 Academic Debate} has the smallest prediction error (0.38 grade point) and the model for \textit{SCI3006 Mathematical Modelling} the largest (1.80 grade point). The mean CV mae weighted by the number of students enrolled in the course is 0.78, the median is 0.78 and the standard deviation is 0.28.

\begin{figure}[htbp]
	\floatconts
	{fig:cv-mae}
	{\caption{Distribution of Cross-Validation Error}}
	{\includegraphics[width=0.5\linewidth]{figures/cv-mae}}
\end{figure}

\subsubsection{Prediction Region for Grade}
\label{sec:warning}
Import from EDM2019

\subsection{The Inductive Confidence Machine (Evgueni)}
\label{sec:icm}

\section{Results}
\label{sec:results}

Make figures

>>>>

% Package to make table with multi rows and columns
\usepackage{multirow}

@@

@@

% start article
% \titlebreak
% \footnote{}
% \textsf

\title[Conformal Prediction for Students' Grades]{Conformal Prediction for Students' Grades in a Course Recommender System}	%\titletag{\thanks{XXX}} % leave empty?

@@

 % Authors with different addresses:
  \author{\Name{Rapha\"{e}l Morsomme} \Email{raphael.morsomme@maastrichtuniversity.nl}\\ % Morsomme\nametag{\thanks{with a note}} % leave this slot empty?
   \addr Zwingelput 4, 6211 KH Maastricht, the Netherlands	
   \AND
   \Name{Evgueni Smirnov} \Email{smirnov@maastrichtuniversity.nl}\\
   \addr Bouillonstraat 10, 6211 LH Maastricht, the Netherlands}

@@

\begin{abstract}
Course selection can be challenging for student of Liberal Arts programs. In particular, due to the highly personalized curricula of liberal arts students, it is often difficult to assess whether or not a particular course is too advanced given their academic background. To assist students of the liberal arts program of the University College Maastricht, \citet{Morsomme.2019} developed a course recommender system that suggests courses whose content matches the student's academic interests, and issues warnings for courses that it deems too advanced.

To issue warnings, the system produces point predictions for the grades that a student will receive in the courses that she/he is considering for the following term. Point predictions are estimated with regression models specific to each course which take into account the academic performance of the student along with the knowledge that she/he has acquired in previous courses. A warning is issued if the predicted grade is a fail.

In this paper, we complement the system's point predictions for grades with prediction intervals constructed using the conformal prediction framework \citep{Vovk.2005}. We use the Inductive Confidence Machine (ICM) \citep{Papadopoulos.2002} with standardized nonconformity scores to construct prediction intervals that are tailored to each student. We find that the prediction intervals constructed with the ICM are valid and that their widths are related to the accuracy of the underlying regression model.
\end{abstract}

\begin{keywords} Conformal Prediction, Recommender System, Course Grade Prediction, Lasso, Education \end{keywords}

% \todo{Rapha\"{e}l: make all figures black and white}

% \todo{Rapha\"{e}l: make sure that all figures have the same font size as the main text}

% \todo{Rapha\"{e}l: use notation Z_k = (x_k, y_k) to simplify description of conformal prediction}

@@

Liberal Arts programs are often characterized by their open curriculum which allows student to tailor their own study program to their academic objectives \citep{Surpatean.2012, Morsomme.2019}. These highly personalized curricula make it difficult for students, academic advisors and course coordinators to assess whether the courses that a student considers taking the following term are too advanced given her/his current academic background or whether she/he has acquired the necessary skills, perhaps through an unusual combination of courses. To alleviate this problem, \citet{Morsomme.2019} developed the Liberal Arts Recommender System (LARS) which suggests to students courses whose content matches their academic interests. In addition, LARS helps student identify courses that are too advanced for them. To accomplish that, the system issues \textit{point predictions} for future grade based on the past academic performance of the student and the skills she/he has acquired in previous courses. A warning is issued when the predicted grade is a fail.

In this paper, we present an application of conformal prediction \citep{Vovk.2005} to complement the current point estimates for future grades of LARS with prediction intervals. For students, the advantage of prediction intervals over point predictions is clear: their information position is improved, thereby enabling them to make better-informed course selection. To avoid the computational costs of transductive conformal prediction, we opt for the lighter Inductive Confidence Machine (ICM) \citep{Papadopoulos.2002}. Furthermore, in order to provide prediction intervals that are tailored to each student, we use a normalized nonconformity measure \citep{Papadopoulos.2011}.

\sectionref{sec:related} presents previous research on grade prediction.
\sectionref{sec:data} introduces the data and \sectionref{sec:crs} briefly describes the existing LARS.
\sectionref{sec:prediction} presents the conformal prediction framework in which we construct the prediction intervals.
\sectionref{sec:experiment} presents the setting and results of the experiment
and \sectionref{sec:conclusion} concludes.

\section{Related Work}
\label{sec:related}

The task of predicting students' course grades prediction has recently received a lot of attention \citep{Polyzou.2016, Houbraken.2017}. The approaches to this problem are regression approaches, classification approaches, and collaborative-filtering approaches. The first two approaches employ general information (secondary education, age, sex etc.), past performance, temporal elements, and contextual information of the students \citep{Bydzovska.2016}. They train a regression/classification model on historical data which is latter used for predicting students' course grades. The collaborative-filtering approaches require only past course grades for future prediction in contrast with the previous two approaches \citep{Sweeney.2015, Houbraken.2017}. They are divided into nearest-neighbor approaches \citep{Bydzovska.2015} and matrix-factorization approaches \citep{Polyzou.2016}. The nearest-neighbor approaches first identify neighbors of a student in terms of study performance and then predict the course grades for this student by aggregating the course grades of the neighbors. The matrix factorization approaches first decompose the student-course data into student and course matrices and then predict student course grades using the product of these matrices.

Although the progress in  student course grade prediction is significant, no research was performed on the problem of estimating the confidence in this type of prediction. As it is stated above we propose to employ conformal prediction for this problem. Our choice is justified by the fact that other approaches for reliable prediction such as version spaces \citep{Smirnov.2004}, meta approaches \citep{Smirnov.2006, Smirnov.2006b}, ROC-isometric approaches \citep{Vanderlooy.2006} are 
inapplicable for regression tasks.

\section{Data for LARS}
\label{sec:data}

\citet{Morsomme.2019} employed two sets of data to develop LARS: student data and course data.

The student data consisted of anonymized course enrollment information. It included the transcripts of the 2,526 students of the liberal arts program between 2008 and 2019 with a total of 79,245 course enrollments. Course enrollments with a missing grade, which indicates that the student either dropped the course or fail the attendance requirement, were removed. \tableref{tab:student-data} presents an example of the student data. Each row contains an anonymized student ID, a course ID, a year and semester, and the obtained grade.

The course data consisted of a corpus of the 490 course descriptions present in the 2018-2019 course catalogues of five departments of Maastricht University: European Studies, University College Maastricht, University College Venlo, Psychology and Science Program. These catalogues contain a one-page description of each course on offer. \tableref{tab:course-data} presents a sample of this textual data for the course \textit{HUM3034 World History} in the tidy format with one row per document-term \citep{Wickham.2014}. The data was processed following common procedures \citep{Meyer.2008}: individual terms were tokenized, stemmed with the Hunspell dictionary and common stop words were removed, as well as numbers between 1 and 1,000 and terms occurring less than 3 times in the corpus.

\begin{table}[hbtp]
	\floatconts
	{tab:student-data}
	{\caption{Example of student data}}	
	{\begin{tabular}{ccccc}
		\toprule
		\bfseries Student ID &\bfseries Course ID &\bfseries Academic Year &\bfseries Period &\bfseries Grade\\
		\midrule
		44940 & CAP3000 & 2009-2010 & 4 & 8.8\\
		37490 & SSC2037 & 2009-2010 & 4 & 8.4\\
		71216 & HUM1003 & 2010-2011 & 4 & 6.8\\
		44212 & SSC2049 & 2010-2011 & 2 & 8.4\\
		85930 & SSC2043 & 2011-2012 & 1 & 4.3\\
		\addlinespace
		14492 & COR1004 & 2012-2013 & 2 & 8.5\\
		34750 & HUM2049 & 2013-2014 & 5 & 6.0\\
		32316 & SSC1001 & 2013-2014 & 1 & 8.5\\
		22092 & SCI1009 & 2014-2015 & 1 & 6.4\\
		19512 & COR1004 & 2016-2017 & 5 & 7.0\\
		\bottomrule
	\end{tabular}}	
\end{table}

\begin{table}[hbtp]
	\floatconts
	{tab:course-data}
	{\caption{Example of course data for the course HUM3034 World History}}	
	{\begin{tabular}{cccc}
			\toprule
			\bfseries Course ID &\bfseries Course Title &\bfseries Department &\bfseries Term\\
			\midrule
			HUM3034 & World History & UCM & understand\\
			HUM3034 & World History & UCM & major\\
			HUM3034 & World History & UCM & issue\\
			HUM3034 & World History & UCM & episode\\
			HUM3034 & World History & UCM & shape\\
			\addlinespace
			HUM3034 & World History & UCM & history\\
			HUM3034 & World History & UCM & mankind\\
			HUM3034 & World History & UCM & focus\\
			HUM3034 & World History & UCM & theme\\
			HUM3034 & World History & UCM & topic\\
			\bottomrule
	\end{tabular}}	
\end{table}

\section{LARS}
\label{sec:crs}
\subsection{Overview}
\label{sec:overview}

LARS is composed of two pillars: course suggestions and warning issuance (see \figureref{fig:flowchart}).

In pillar 1, a topic model of the courses is fitted to the course data using the Latent Dirichlet Allocation model \citep{Blei.2003}. A topic model represents a topic as a mixture of words and a document as a mixture of topics. The key words selected by the student are then mapped to the vocabulary of the topic model to represent her/his academic interests. Finally, the system matches the student's academic interests to the content of the courses as represented by the topic model to identify courses of interest to the student.

In pillar 2 for warning issuance, a model of each student is first created which contains information about the academic performance of the student (derived from the student data) and the expertise in specific topics (derived from the topic model) that she/he has acquired in previous courses. A regression model for point prediction of the grades that takes the student model as input is then fitted for each course. LARS uses these models to predict the grade that the student will obtain in the courses that she/he is considering for the following term and issues a warning when the predicted grade is a fail.

\begin{figure}[htbp]
	\floatconts
	{fig:flowchart}
	{\caption{Original LARS (black) and our contribution (red)}}
	{\includegraphics[width=0.5\linewidth]{figures/flowchart}}
\end{figure}

\subsection{Pillar 1: Course Suggestion}

\subsubsection{Topic Model of the Courses}
\label{sec:tm}

\citet{Morsomme.2019} fitted a topic model to the course data using the Latent Dirichlet Allocation generative model \citep{Blei.2003} and the Gibbs sampling algorithm \citep{Phan.2008}. The LDA conceptualizes topics as a probability distribution over the vocabulary of the corpus, and document as a set of words, each drawn from a probability distribution over topics specific to that document. The term \textit{Dirichlet} comes from the fact that the word distribution $\beta_{t}$ of topic $t$ is generated from a Dirichlet distribution $\beta_{t} \sim Dirichlet(\delta)$ and the topic distribution $\theta_{d}$ for document $d$ is generated from a Dirichlet distribution $\theta_{d} \sim Dirichlet(\alpha)$ where $\delta$ and $\alpha$ act as hyper-parameters determining how concentrated the distributions of words in topics and the distributions of topics in documents are.

The authors of LARS followed \citet{Phan.2008} who use a Gibbs sampler to learn the distributions $\beta$ and $\theta$ of each topic and document, and \citet{Griffiths.2004} who select the number of topics yielding the best model with respect to the log-likelihood. The selected topic model contains 65 topics (see \figureref{fig:ll}) and consists of a term distribution for each topic indicating the importance of each term of the corpus in the topic and a topic distribution for each course indicating the importance of each topic in the course. \figureref{fig:gamma} shows the main topics of the core course \textit{COR1004 Political Philosophy} and \figureref{fig:beta} presents the terms that characterize topic 4 and topic 19, the main two topics of the course. We can see that the topics are easy to interpret and that the content of the course, that is, its topic distribution, corresponds to what we would expect from a course on political philosophy.

\begin{figure}[htbp]
	\floatconts
	{fig:ll}
	{\caption{Maximum likelihood model selection: the model with 65 topics is selected.}}
	{\includegraphics[width=0.5\linewidth]{figures/model-selection}}
\end{figure}

\begin{figure}[htbp]
	\floatconts
	{fig:gamma}
	{\caption{Topic distribution in the course COR1004 Political Philosophy.}}
	{\includegraphics[width=0.5\linewidth]{figures/gamma-COR1004}}
\end{figure}

\begin{figure}[htbp]
	\floatconts
	{fig:beta}
	{\caption{Word distribution in the main two topics of COR1004 Political Philosophy. Topic 4 corresponds to international politics and Topic 19 to philosophy.}}
	{\includegraphics[width=0.75\linewidth]{figures/beta}}
\end{figure}

\subsubsection{Model of a Student's Academic Interests}

\citet{Morsomme.2019} employed the topic model to estimate the academic interests of a student from the key words that she/he enters into the system. The student's academic interest $AI_t$ in topic $t$ simply corresponds to the sum of the selected key words' importance in topic $t$ as determined by the topic model, that is,
$$ AI_t = \sum_{i \in I^*} \beta_{t,i}, \;\;\;  \mbox{for} \;\;\;  t=1, \cdots, n,$$
where $I^*$ is the set of key words selected by the student, $\beta_{t,i}$ corresponds to the importance of term $i$ in topic $t$ and $n$ is the number of topics present in the model (in this case $n=65$). The vector $AI = (AI_1, \cdots, AI_{n})^T$ therefore represents the academic interests of the student.

\subsubsection{Course Matching and Suggestion}

LARS uses the Kullback-Leibler (KL) divergence \citep{Kullback.1951} to identify the courses whose content best matche the academic interests of the student.
Letting $P$ and $Q$ be two discrete probability distribution defined on the same probability space, the KL divergence between $P$ and $Q$ is defined as

$$ D_{\mathrm{KL}} \left( P || Q \right) =  - \sum_{x \in X} P(x) \log\left(\dfrac{Q(x)}{P(x)}\right)$$

\noindent and measures how different the probability distribution $P$ is from the reference probability distribution $Q$.
LARS suggests to the students the $n$ courses whose topic distribution $\theta$ has the smallest KL divergence to her/his normalized academic interests $AI^* = \dfrac{AI}{\lvert AI \rvert}$.

\subsection{Pillar 2: Warning Issuance}
\subsubsection{Student Model}
\label{sec:sm}
The student model consists of two elements: academic performance and topic-specific expertise. Academic performance corresponds to the student's general GPA (grade point average) as well as her/his GPA in humanities, natural sciences, social sciences, skills and projects. These are derived from the students' transcripts in a straightforward way. Topic-specific expertise corresponds to the amount of knowledge that the student has acquired  in previous courses in each of the topics present in the topic model. \citet{Morsomme.2019}  posited that, when students take a course, they acquire knowledge about its content and that the amount of knowledge that they acquire is proportional to the obtained grade; that is, they assume that students who obtain 10/10 in a course acquire all the knowledge related to its content while those who obtain 5/10 only acquire half of it. The content of a course is determined by its topic distribution in the topic model and the grades are retrieved from the student's transcript. Furthermore, they assume that the knowledge acquired in different courses simply accumulates. Hence, if a student has taken $n$ courses and $g_{i}$ corresponds to her/his grade in course $i$, for $i = 1, \cdots, n$, then her/his expertise $exp_t$ in topic $t$ corresponds to

\begin{equation}
\label{eq:expertise}
exp_t = \sum_{i= 1}^{n} g_i \theta_{i,t}
\end{equation}

\noindent where $\theta_{i,t}$ corresponds to the importance of topic $t$ in course $i$ as determined by the topic model. \tableref{tab:toy} and \figureref{fig:toy-expertise} present a toy example of the contribution of three individual courses toward a student's expertise in five topics. \tableref{tab:toy-gamma} and \tableref{tab:toy-grade} respectively show the topic distribution in each course as estimated by some topic model and the grades obtained by the student which are retrieved from her/his transcript. \tableref{tab:toy-contribution} uses \equationref{eq:expertise} to estimate the contribution of each course toward the student's topic expertise. \figureref{fig:toy-expertise} offers a graphical illustration of \tableref{tab:toy-contribution}.

\begin{table}[htbp]
	\floatconts
	{tab:toy}
	{\caption{Toy example of the contribution of individual courses toward a student's topic expertise.}}
	{%
		\subtable[Topic distribution $\theta$]{%
			\label{tab:toy-gamma}%
			\begin{tabular}{cccccc}
				\toprule
				\bfseries Course & \bfseries Topic 1 & \bfseries Topic 2 & \bfseries Topic 3 & \bfseries Topic 4 & \bfseries Topic 5\\
				\midrule
				Course 1 & 0.0 & 0.7 & 0.2 & 0.1 & 0.0\\
				Course 2 & 0.2 & 0.2 & 0.2 & 0.2 & 0.2\\
				Course 3 & 0.0 & 0.4 & 0.2 & 0.1 & 0.2\\
				\bottomrule
			\end{tabular}
		}\qquad
		\subtable[Transcript]{%
			\label{tab:toy-grade}%
			\begin{tabular}{cc}
				\toprule
				\bfseries Course & \bfseries Grade\\
				\midrule
				Course 1 & 6/10\\
				Course 2 & 9/10\\
				Course 3 & 2/10\\
				\bottomrule
			\end{tabular}
		}\qquad
			\subtable[Course contribution toward a student's topic expertise]{%
			\label{tab:toy-contribution}%
			\begin{tabular}{cccccc}
				\toprule
				\bfseries Course & \bfseries Topic 1 & \bfseries Topic 2 & \bfseries Topic 3 & \bfseries Topic 4 & \bfseries Topic 5\\
				\midrule
				Course 1 & 0.00 & 0.42 & 0.12 & 0.06 & 0.00\\
				Course 2 & 0.18 & 0.18 & 0.18 & 0.18 & 0.18\\
				Course 3 & 0.00 & 0.08 & 0.04 & 0.02 & 0.04\\
				\midrule
				\bfseries Total & 0.18 & 0.68 & 0.34 & 0.26 & 0.22\\
				\bottomrule
			\end{tabular}
		}
	}
\end{table}

\begin{figure}[htbp]
	\floatconts
	{fig:toy-expertise}
	{\caption{Toy example of the contribution of individual courses toward a student's topic expertise.}}
	{\includegraphics[width=0.5\linewidth]{figures/toy-expertise}}
\end{figure}

\subsubsection{Point Prediction for Grade and Warning Issuance}

To issue warnings, LARS produces point estimates for future grades. To accomplish this, it uses regression models.  LARS separately fits a sparse linear regression model for grade prediction to each of the 132 courses currently offered at the University College Maastricht with at least 20 student enrollments since 2008. The input to the models consists of the 71 variables stored in the student model: 6 GPAs (1 general and 5 discipline-specific) and the level of expertise in the 65 topics of the topic model. The regression models output a point estimate for the grade. Note that each model is trained only on the data of the students enrolled in the associated course. Since the number of predictors is relatively large, the models are regularized with the Lasso penalty \citep{Tibshirani.1996} and the value of the Lasso tuning parameter $\lambda$ is chosen via cross-validation (CV). \figureref{fig:cv-mae} shows the distribution of the CV mean absolute error (mae) for the 132 prediction models. The model for the course \textit{PRO2004 Academic Debate} has the smallest prediction error (0.38 grade point) and the model for \textit{SCI3006 Mathematical Modelling} the largest (1.80 grade point). The mean CV mae weighted by the number of students enrolled in the course is 0.78, the median is 0.78 and the standard deviation is 0.28.

% use cross-validation to find the value of the Lasso tuning parameter $lambda$ that minimizes the CV mean absolute error (mae), a more robust loss function than the squared absolute error \citep{Hastie.2009}

\begin{figure}[htbp]
	\floatconts
	{fig:cv-mae}
	{\caption{Distribution of cross-validation error}}
	{\includegraphics[width=0.5\linewidth]{figures/cv-mae}}
\end{figure}

In practice, the student selects the courses that she/he is considering for the following term and the system uses the regression models of these courses to provide point predictions for the future grades. A warning is issued when a predicted grade is a fail.

We desire the following three requirements for pillar 2 Warning Issuance of LARS: accuracy, sparsity and transparency of the regression models for grade prediction. First, the grade predictions must be accurate so that students base their course selection on sound information. Second, we want the regression models to be sparse so that we can identify which topics are important to master in order to perform well in a given course. Such information would be extremely useful to course coordinators and curriculum managers alike. Third, in order to be transparent, grade predictions must be accompanied by an indication of their own accuracy. The first two requirements are fulfilled by the Lasso model \citep{Tibshirani.1996}, but the third requirement is not satisfied by LARS's current \textit{point predictions} for future grades. To fulfill the requirement for transparency, we propose to complement the existing point predictions of the system with \textit{prediction intervals}. In the following section, we use the conformal prediction framework to build prediction intervals for future grades that are tailored to each student.

\section{Regression with Prediction Interval}
\label{sec:prediction}
